---
title: "Project"
author: "Kate O'Rourke"
date: "2024-10-29"
output: pdf_document
---

# Load Libraries
```{r}
library(tidyverse)
library(haven)
library(gtsummary)
```

# Read in .XPT data file - KO
```{r}
CDC_2023 <- read_xpt('LLCP2023.XPT')
```

#### Remove unused columns - KO
```{r}
CDC_2023_subset = subset(CDC_2023, 
                         select = c(EDUCA, EMPLOY1, INCOME3, WEIGHT2, HEIGHT3, SEXVAR, PHYSHLTH, MENTHLTH, PERSDOC3,
                                               MEDCOST1, SMOKE100, DRNK3GE5, EXERANY2, BPHIGH6, TOLDHI3, CVDINFR4,
                                               CVDCRHD4, CVDSTRK3, CHCKDNY2, DIABETE4, DIABTYPE, `_AGE80`)
                         )
```

#### Show NA values - KO
```{r}
colSums(is.na(CDC_2023_subset))
```

### Convert columns (except for WEIGHT2, HEIGHT3, DRNK3GE5, PHYSHLTH, MENTHLTH, AGE) to factors - KO
```{r}
col_names <- colnames(CDC_2023_subset)
col_names <- col_names[-c(4, 5, 7, 8, 12, 22)] # Columns 4 (WEIGHT2), 5 (HEIGHT3), 7 (PHYSHLTH), 8 (MENTHLTH), 12 (DRNK3GE5) and 22 (_AGE80)
CDC_2023_subset[,col_names] <- lapply(CDC_2023_subset[,col_names] , factor)
```

### Get columns summaries for factor data - KO
```{r}
CDC_2023_subset %>%
  tbl_summary()
```


# Data Cleaning

## Filter out rows with metric measurements (e.g., codes like 9061-9998 for metric values) - SG
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  filter(
    !(HEIGHT3 >= 9000),
    !(WEIGHT2 >=9000),
    WEIGHT2 != 7777,       # Filter out any values where person refused or didn't know weight
    HEIGHT3 != 7777
    )

# Extract feet and inches assuming HEIGHT3 values are in a format like '511' (5 feet, 11 inches)
CDC_2023_subset <- CDC_2023_subset %>%
  mutate(
    feet = as.numeric(substr(HEIGHT3, 1, 1)),  # Extract first digit as feet
    inches = as.numeric(substr(HEIGHT3, 2, 3)),  # Extract next two as inches
    height_inches = feet * 12 + inches  # Convert total height to inches
  ) %>%
  select(-feet, -inches) 
```

## Calculate BMI and add it as a new column - SG
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  mutate(BMI = (WEIGHT2 * 703) / (height_inches^2))
```

## Filter to include individuals with type 2 diabetes, prediabetes, or no diabetes - SG
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  filter(
    (DIABETE4 == 1 & DIABTYPE == 2) |  # Keep those with type 2 diabetes
    (DIABETE4 == 4) |                  # Keep those with prediabetes
    (DIABETE4 == 3)                    # Keep those with no diabetes
  )
```

## Remove rows with missing or refused data codes - SG
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  filter(
    EDUCA != 9,             # Refused education level
    EMPLOY1 != 9,           # Refused employment status
    INCOME3 != 77 & INCOME3 != 99,    # Don't know or refused income level
    PHYSHLTH != 77 & PHYSHLTH != 99,  # Don't know or refused physical health
    MENTHLTH != 77 & MENTHLTH != 99,  # Don't know or refused mental health
    PERSDOC3 != 7 & PERSDOC3 != 9,    # Don't know or refused personal doctor info
    MEDCOST1 != 7 & MEDCOST1 != 9,    # Don't know or refused medical cost
    SMOKE100 != 7 & SMOKE100 != 9,    # Don't know or refused smoking info
    DRNK3GE5 != 77 & DRNK3GE5 != 99,  # Don't know or refused binge drinking info
    EXERANY2 != 7 & EXERANY2 != 9,    # Don't know or refused exercise info
    BPHIGH6 != 7 & BPHIGH6 != 9,      # Don't know or refused high blood pressure
    TOLDHI3 != 7 & TOLDHI3 != 9,      # Don't know or refused high cholesterol
    CVDINFR4 != 7 & CVDINFR4 != 9,    # Don't know or refused heart attack history
    CVDCRHD4 != 7 & CVDCRHD4 != 9,    # Don't know or refused coronary heart disease
    CVDSTRK3 != 7 & CVDSTRK3 != 9,    # Don't know or refused stroke history
    CHCKDNY2 != 7 & CHCKDNY2 != 9,    # Don't know or refused kidney disease
    DIABETE4 != 7 & DIABETE4 != 9,    # Don't know or refused diabetes status
    `_AGE80` != 7 & `_AGE80` != 9     # Don't know or refused age
  )

```

## Simplify variables - KO
```{r}
# EMPLOY1 is simplified into 1 (employed: employed for wages, self-employed), 2 (out of labor force: homemaker, student, retired), 3 (unemployed: our of work for 1 year or more, out of work for less than 1 year, unable to work)
CDC_2023_subset$EMPLOY1 <- as.numeric(CDC_2023_subset$EMPLOY1)
CDC_2023_subset$EMPLOY1 <- replace(CDC_2023_subset$EMPLOY1, CDC_2023_subset$EMPLOY1 == 1 | CDC_2023_subset$EMPLOY1 == 2, 1) 
CDC_2023_subset$EMPLOY1 <- replace(CDC_2023_subset$EMPLOY1, CDC_2023_subset$EMPLOY1 == 5 | CDC_2023_subset$EMPLOY1 == 6 | CDC_2023_subset$EMPLOY1 == 7, 2) 
CDC_2023_subset$EMPLOY1 <- replace(CDC_2023_subset$EMPLOY1, CDC_2023_subset$EMPLOY1 == 3 | CDC_2023_subset$EMPLOY1 == 4 | CDC_2023_subset$EMPLOY1 == 8, 3) 
CDC_2023_subset$EMPLOY1 <- as.factor(CDC_2023_subset$EMPLOY1)

# PHYSHLTH is simplified to 1 (zero days when physical health was not good - 88), 2 (1 to 13 days when physical health is not good), 3 (14+ days when physical health is not good)
CDC_2023_subset$PHYSHLTH <- replace(CDC_2023_subset$PHYSHLTH, CDC_2023_subset$PHYSHLTH == 88, 0)
#CDC_2023_subset$PHYSHLTH <- replace(CDC_2023_subset$PHYSHLTH, CDC_2023_subset$PHYSHLTH >= 1 & #CDC_2023_subset$PHYSHLTH < 14, 1) 
#CDC_2023_subset$PHYSHLTH <- replace(CDC_2023_subset$PHYSHLTH, CDC_2023_subset$PHYSHLTH >= 14, 2) 
#CDC_2023_subset$PHYSHLTH <- as.factor(CDC_2023_subset$PHYSHLTH)

# MENTHLTH is simplified to 1 (zero days when mental health was not good - 88), 2 (1 to 13 days when mental health is not good), 3 (14+ days when mental health is not good)
CDC_2023_subset$MENTHLTH <- replace(CDC_2023_subset$MENTHLTH, CDC_2023_subset$MENTHLTH == 88, 0) 
#CDC_2023_subset$MENTHLTH <- replace(CDC_2023_subset$MENTHLTH, CDC_2023_subset$MENTHLTH >= 1 & CDC_2023_subset$MENTHLTH < 14, 1)
#CDC_2023_subset$MENTHLTH <- replace(CDC_2023_subset$MENTHLTH, CDC_2023_subset$MENTHLTH >= 14, 2) 
#CDC_2023_subset$MENTHLTH <- as.factor(CDC_2023_subset$MENTHLTH)

# PERSDOC3 will be simplified to combine 1 and 2 to indicate that the person has at least one personal doctor and 3 will be converted to 2
CDC_2023_subset$PERSDOC3 <- as.numeric(CDC_2023_subset$PERSDOC3)
CDC_2023_subset$PERSDOC3 <- replace(CDC_2023_subset$PERSDOC3, CDC_2023_subset$PERSDOC3 == 1 & CDC_2023_subset$PERSDOC3 == 2, 1) 
CDC_2023_subset$PERSDOC3 <- replace(CDC_2023_subset$PERSDOC3, CDC_2023_subset$PERSDOC3 == 3, 2)
CDC_2023_subset$PERSDOC3 <- as.factor(CDC_2023_subset$PERSDOC3)

# DRNK3GE5 will be simplified to indicate that a person has been binge drinking at least 1 time in last 30 days with 1 and no drinks will be indicated by 0 (originally indicated by 88)
CDC_2023_subset$DRNK3GE5 <- replace(CDC_2023_subset$DRNK3GE5, CDC_2023_subset$DRNK3GE5 == 88, 0)
#CDC_2023_subset$DRNK3GE5 <- replace(CDC_2023_subset$DRNK3GE5, CDC_2023_subset$DRNK3GE5 >= 1, 1) 
#CDC_2023_subset$DRNK3GE5 <- as.factor(CDC_2023_subset$DRNK3GE5)

# DIABETE4 will be simplified so 0 indicates the individual does not have diabetes (3) and 1 indicating the individual has type 2 diabetes or pre-diabetes or borderline diabetes (1 and 2 for DIABTYPE or 4)
CDC_2023_subset$DIABETE4 <- as.numeric(CDC_2023_subset$DIABETE4)
CDC_2023_subset$DIABETE4 <- replace(CDC_2023_subset$DIABETE4, CDC_2023_subset$DIABETE4 == 3, 0)
CDC_2023_subset$DIABETE4 <- replace(CDC_2023_subset$DIABETE4, CDC_2023_subset$DIABETE4 == 1 | CDC_2023_subset$DIABETE4 == 4, 1)
CDC_2023_subset$DIABETE4 <- as.factor(CDC_2023_subset$DIABETE4)

# BPHIGH6 will be simplified so 1 indicates one has had high or borderline high or pre-hypertensive or elevated blood pressure
CDC_2023_subset$BPHIGH6 <- as.numeric(CDC_2023_subset$BPHIGH6)
CDC_2023_subset$BPHIGH6 <- replace(CDC_2023_subset$BPHIGH6, CDC_2023_subset$BPHIGH6 == 2 | CDC_2023_subset$BPHIGH6 == 3, 2)
CDC_2023_subset$BPHIGH6 <- replace(CDC_2023_subset$BPHIGH6, CDC_2023_subset$BPHIGH6 == 1 | CDC_2023_subset$BPHIGH6 == 4, 1)
CDC_2023_subset$BPHIGH6 <- as.factor(CDC_2023_subset$BPHIGH6)
```

## Export new data file as cvs file - KO
```{r}
write.csv(CDC_2023_subset, "CDC_2023_cleaned.csv")
```
