---
title: "Logistic Regression All Variables Model"
author: "Kate O'Rourke"
date: "2024-11-21"
output: pdf_document
---

## Libraries
```{r}
library(dplyr)
library(caret) # used to split dataset
library(readr)
library(ggplot2)
library(pROC)
library(tidymodels) # Random Forest
```

# Preprocessing

## Read in dataset - KO
```{r}
CDC_2023_subset <- read_csv("CDC_2023_cleaned.csv")
```

## Remove unused columns - KO
```{r}
CDC_2023_subset <-  subset(CDC_2023_subset, select = -c(1, 5:6, 22, 24) )
```

## Convert columns (except for EDUCA, EMPLOY1, INCOME3, PHYSHLTH, MENTHLTH, DRNK3GE5, _AGE80, BMI) to factors - KO
```{r}
col_names <- colnames(CDC_2023_subset)
col_names <- col_names[-c(1:3, 5:6, 10, 19, 20)] # Column 1 (EDUCA), 2 (EMPLOY1), 3 (INCOME3), 5 (PHYSHLTH), 6 (MENTHLTH), 10 (DRNK3GE5), 19 (_AGE80), 20 (BMI)
CDC_2023_subset[,col_names] <- lapply(CDC_2023_subset[,col_names] , factor)
```

## Function to normalize BMI column - KO
```{r}
# Function to normalize specific columns based on dataset and columns specified
normalize <- function(dataset, columns) {
  scale(dataset[, columns],
  center = apply(dataset[, columns], 2, mean),
  scale = apply(dataset[, columns], 2, sd)
  )
}
```

## Split dataset into training and test sets - KO
Dataset will be split into 77% training and 23% testing.
```{r}
training <- createDataPartition(CDC_2023_subset$DIABETE4,
                                         p = 0.77,
                                         list = FALSE,
                                         times = 1)

CDC_2023_training <- CDC_2023_subset[training, ]
CDC_2023_test <- CDC_2023_subset[-training, ]
```

## Normalize Numeric Columns in training and test sets - KO
```{r}
columns = c(1:3, 5:6, 10, 19, 20)
CDC_2023_training[, columns] <- normalize(CDC_2023_training, columns)
CDC_2023_test[, columns] <- normalize(CDC_2023_test, columns)
```

## Logistic Regression

```{r}
diabetes_recipe <- 
  recipe(
    DIABETE4 ~ ., 
    data = CDC_2023_training
  ) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  prep(training = CDC_2023_training)

logreg_fit <- 
  logistic_reg() %>%
  set_engine("glm") %>%
  fit(DIABETE4 ~ ., data = bake(diabetes_recipe, new_data = CDC_2023_training))
logreg_fit

test_baked <- bake(diabetes_recipe, new_data = CDC_2023_test, all_predictors())

test_results <- 
  CDC_2023_test %>%
  dplyr::select(DIABETE4) %>%
  bind_cols(
    predict(logreg_fit, new_data = test_baked, type = "prob") %>%
      dplyr::select(p_1 = .pred_1)
  )

################################### plot ROC curve ###################################

roc_data <- data.frame(threshold=seq(1,0,-0.01), fpr=0, tpr=0)
for (i in roc_data$threshold) {
  
  over_threshold <- test_results[test_results$p_1 >= i, ]
  
  fpr <- sum(over_threshold$DIABETE4==0)/sum(test_results$DIABETE4==0)
  roc_data[roc_data$threshold==i, "fpr"] <- fpr
  
  tpr <- sum(over_threshold$DIABETE4==1)/sum(test_results$DIABETE4==1)
  roc_data[roc_data$threshold==i, "tpr"] <- tpr
  
}

ggplot() +
  geom_line(data = roc_data, aes(x = fpr, y = tpr, color = threshold), linewidth = 2) +
  scale_color_gradientn(colors = rainbow(3)) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(data = roc_data[seq(1, 101, 10), ], aes(x = fpr, y = tpr)) +
  geom_text(data = roc_data[seq(1, 101, 10), ],
            aes(x = fpr, y = tpr, label = threshold, hjust = 1.2, vjust = -0.2))

###################################          AUC           ###################################

roc_object <- roc( CDC_2023_test$DIABETE4, test_results$p_1)
auc(roc_object)

###############################          ACCURACY          ###################################

pred <- rep(0, 2532)
pred[test_results$p_1 > 0.5] <- 1
table(pred, CDC_2023_test$DIABETE4)
paste("Accuracy = ", mean(pred == CDC_2023_test$DIABETE4))

################################### plot calibration curve ###################################

calibration_data <- data.frame(bin_midpoint=seq(0.05,0.95,0.1),
                               observed_event_percentage=0)
for (i in seq(0.05,0.95,0.1)) {
  
  in_interval <- test_results[test_results$p_1 >= (i-0.05) & test_results$p_1 <= (i+0.05), ]
  oep <- nrow(in_interval[in_interval$DIABETE4==1, ])/nrow(in_interval)
  calibration_data[calibration_data$bin_midpoint==i, "observed_event_percentage"] <- oep
  
}

ggplot(data = calibration_data, aes(x = bin_midpoint, y = observed_event_percentage)) +
  geom_line(linewidth = 1) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(size = 2) +
  geom_text(aes(label = bin_midpoint), hjust = 0.75, vjust = -0.5)
```