---
title: "Preprocessing and Feature Engineering"
author: "Kate O'Rourke"
date: "2024-11-14"
output: pdf_document
---
# General Preprocessing Steps for Feature Engineering

# Libraries
```{r}
library(dplyr)
library(caret) # used to split dataset
library(readr)
```

# Read in dataset - KO
```{r}
CDC_2023_subset <- read_csv("CDC_2023_cleaned.csv")
```

# Remove unused columns - KO
```{r}
CDC_2023_subset = subset(CDC_2023_subset, select = -c(1, 5:6, 22, 23) )
```

# Convert columns (except for BMI) to factors - KO
```{r}
col_names <- colnames(CDC_2023_subset)
col_names <- col_names[-19] # Column 19 (BMI)
CDC_2023_subset[,col_names] <- lapply(CDC_2023_subset[,col_names] , factor)
```

# Function to normalize BMI column - KO
```{r}
# Function to normalize specific columns based on dataset and columns specified
normalize <- function(dataset, columns) {
  dataset[, columns] <- scale(dataset[, columns],
                              center = apply(dataset[, columns], 2, mean),
                              scale = apply(dataset[, columns], 2, sd)
                              )
}
```

# Calculate test and training split - KO (function sourced from Katherine S. Geist)
```{r}
## Written by Katherine S. Geist, PhD
## Merrimack College, Massachusetts
## Please do not distribute without attribution
## https://github.com/ksgeist

calcSplitRatio <- function(p = NA, df) {
  ## @p  = the number of parameters. by default, if none are provided, the number of columns (predictors) in the dataset are used
  ## @df = the dataframe that will be used for the analysis
  
  ## If the number of parameters isn't supplied, set it to the number of features minus 1 for the target
  if(is.na(p)) {
    p <- ncol(df) -1   ## COMMENT HERE
  }
  
  ## Calculate the ideal number of testing set
  test_N <- (1/sqrt(p))*nrow(df)
  ## Turn that into a testing proportion
  test_prop <- round((1/sqrt(p))*nrow(df)/nrow(df), 2)
  ## And find the training proportion
  train_prop <- 1-test_prop
  
  ## Tell us the results!
  print(paste0("The ideal split ratio is ", train_prop, ":", test_prop, " training:testing"))
  
  ## Return the size of the training set
  return(train_prop)
}

calcSplitRatio(p=NA, CDC_2023_subset)
```
Dataset will be split into 76% training and 24% testing.

# Split dataset into training and test sets - KO
```{r}
training <- createDataPartition(CDC_2023_subset$DIABETE4,
                                         p = 0.76,
                                         list = FALSE,
                                         times = 1)

CDC_2023_training <- CDC_2023_subset[training, ]
CDC_2023_test <- CDC_2023_subset[-training, ]
```

# Normalize BMI in training and test sets - KO
```{r}
columns = 19
normalize(CDC_2023_training, columns)
normalize(CDC_2023_test, columns)
```

# Check proportions of DIABETE4 in training and test sets - KO
```{r}
prop.table(table(CDC_2023_training$DIABETE4))
prop.table(table(CDC_2023_test$DIABETE4))
```