---
title: "Project"
author: "Kate O'Rourke"
date: "2024-10-29"
output: pdf_document
---

```{r}
library(tidyverse)
library(haven)
```

```{r}
CDC_2023 <- read_xpt('LLCP2023.XPT')
```

#### Remove unused columns
```{r}
CDC_2023_subset = subset(CDC_2023, 
                         select = c(EDUCA, EMPLOY1, INCOME3, WEIGHT2, HEIGHT3, SEXVAR, PHYSHLTH, MENTHLTH, PERSDOC3,
                                               MEDCOST1, SMOKE100, DRNK3GE5, EXERANY2, BPHIGH6, TOLDHI3, CVDINFR4,
                                               CVDCRHD4, CVDSTRK3, CHCKDNY2, DIABETE4, DIABTYPE)
                         )
```

```{r}
colSums(is.na(CDC_2023_subset))
```

### Convert columns (except for WEIGHT2, HEIGHT3, DRNK3GE5, PHYSHLTH, MENTHLTH) to factors
```{r}
col_names <- colnames(CDC_2023_subset)
col_names <- col_names[-c(4, 5, 7, 8, 12)] # Columns 4 (WEIGHT2), 5 (HEIGHT3), 7 (PHYSHLTH), 8 (MENTHLTH) and 12 (DRNK3GE5)
CDC_2023_subset[,col_names] <- lapply(CDC_2023_subset[,col_names] , factor)
```

### Get columns summaries for factor data
```{r}
library(gtsummary)
```

```{r}
CDC_2023_subset %>%
  tbl_summary()
```


# Data Cleaning

## Filter out rows with metric measurements (e.g., codes like 9061-9998 for metric values)
```{r}

CDC_2023_subset <- CDC_2023_subset %>%
  filter(
    !(HEIGHT3 >= 9000),
    !(WEIGHT2 >=9000),
    WEIGHT2 != 7777,       # Filter out any values where person refused or didn't know weight
    HEIGHT3 != 7777
    )

# Extract feet and inches assuming HEIGHT3 values are in a format like '511' (5 feet, 11 inches)
CDC_2023_subset <- CDC_2023_subset %>%
  mutate(
    feet = as.numeric(substr(HEIGHT3, 1, 1)),  # Extract first digit as feet
    inches = as.numeric(substr(HEIGHT3, 2, 3)),  # Extract next two as inches
    height_inches = feet * 12 + inches  # Convert total height to inches
  ) %>%
  select(-feet, -inches) 
```

## Calculate BMI and add it as a new column
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  mutate(BMI = (WEIGHT2 * 703) / (height_inches^2))
```

## Filter to include individuals with type 2 diabetes, prediabetes, or no diabetes
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  filter(
    (DIABETE4 == 1 & DIABTYPE == 2) |  # Keep those with type 2 diabetes
    (DIABETE4 == 4) |                  # Keep those with prediabetes
    (DIABETE4 == 3)                    # Keep those with no diabetes
  )
```

## Remove rows with missing or refused data codes
```{r}
CDC_2023_subset <- CDC_2023_subset %>%
  filter(
    EDUCA != 9,             # Refused education level
    EMPLOY1 != 9,           # Refused employment status
    INCOME3 != 77 & INCOME3 != 99,    # Don't know or refused income level
    PHYSHLTH != 77 & PHYSHLTH != 99,  # Don't know or refused physical health
    MENTHLTH != 77 & MENTHLTH != 99,  # Don't know or refused mental health
    PERSDOC3 != 7 & PERSDOC3 != 9,    # Don't know or refused personal doctor info
    MEDCOST1 != 7 & MEDCOST1 != 9,    # Don't know or refused medical cost
    SMOKE100 != 7 & SMOKE100 != 9,    # Don't know or refused smoking info
    DRNK3GE5 != 77 & DRNK3GE5 != 99,  # Don't know or refused binge drinking info
    EXERANY2 != 7 & EXERANY2 != 9,    # Don't know or refused exercise info
    BPHIGH6 != 7 & BPHIGH6 != 9,      # Don't know or refused high blood pressure
    TOLDHI3 != 7 & TOLDHI3 != 9,      # Don't know or refused high cholesterol
    CVDINFR4 != 7 & CVDINFR4 != 9,    # Don't know or refused heart attack history
    CVDCRHD4 != 7 & CVDCRHD4 != 9,    # Don't know or refused coronary heart disease
    CVDSTRK3 != 7 & CVDSTRK3 != 9,    # Don't know or refused stroke history
    CHCKDNY2 != 7 & CHCKDNY2 != 9,    # Don't know or refused kidney disease
    DIABETE4 != 7 & DIABETE4 != 9     # Don't know or refused diabetes status
  )

```

## Simplify variables
```{r}
# EMPLOY1 is simplified into 1 (employed: employed for wages, self-employed), 2 (out of labor force: homemaker, student, retired), 3 (unemployed: our of work for 1 year or more, out of work for less than 1 year, unable to work)
CDC_2023_subset$EMPLOY1 <- as.numeric(CDC_2023_subset$EMPLOY1)
CDC_2023_subset$EMPLOY1 <- replace(CDC_2023_subset$EMPLOY1, CDC_2023_subset$EMPLOY1 == 1 | CDC_2023_subset$EMPLOY1 == 2, 1) 
CDC_2023_subset$EMPLOY1 <- replace(CDC_2023_subset$EMPLOY1, CDC_2023_subset$EMPLOY1 == 5 | CDC_2023_subset$EMPLOY1 == 6 | CDC_2023_subset$EMPLOY1 == 7, 2) 
CDC_2023_subset$EMPLOY1 <- replace(CDC_2023_subset$EMPLOY1, CDC_2023_subset$EMPLOY1 == 3 | CDC_2023_subset$EMPLOY1 == 4 | CDC_2023_subset$EMPLOY1 == 8, 3) 
CDC_2023_subset$EMPLOY1 <- as.factor(CDC_2023_subset$EMPLOY1)

# PHYSHLTH is simplified to 1 (zero days when physical health was not good - 88), 2 (1 to 13 days when physical health is not good), 3 (14+ days when physical health is not good)
CDC_2023_subset$PHYSHLTH <- replace(CDC_2023_subset$PHYSHLTH, CDC_2023_subset$PHYSHLTH == 88, 0)
CDC_2023_subset$PHYSHLTH <- replace(CDC_2023_subset$PHYSHLTH, CDC_2023_subset$PHYSHLTH >= 1 & CDC_2023_subset$PHYSHLTH < 14, 1) 
CDC_2023_subset$PHYSHLTH <- replace(CDC_2023_subset$PHYSHLTH, CDC_2023_subset$PHYSHLTH >= 14, 2) 
CDC_2023_subset$PHYSHLTH <- as.factor(CDC_2023_subset$PHYSHLTH)

# MENTHLTH is simplified to 1 (zero days when mental health was not good - 88), 2 (1 to 13 days when mental health is not good), 3 (14+ days when mental health is not good)
CDC_2023_subset$MENTHLTH <- replace(CDC_2023_subset$MENTHLTH, CDC_2023_subset$MENTHLTH == 88, 0) 
CDC_2023_subset$MENTHLTH <- replace(CDC_2023_subset$MENTHLTH, CDC_2023_subset$MENTHLTH >= 1 & CDC_2023_subset$MENTHLTH < 14, 1)
CDC_2023_subset$MENTHLTH <- replace(CDC_2023_subset$MENTHLTH, CDC_2023_subset$MENTHLTH >= 14, 2) 
CDC_2023_subset$MENTHLTH <- as.factor(CDC_2023_subset$MENTHLTH)

# PERSDOC3 will be simplified to combine 1 and 2 to indicate that the person has at least one personal doctor and 3 will be converted to 2
CDC_2023_subset$PERSDOC3 <- as.numeric(CDC_2023_subset$PERSDOC3)
CDC_2023_subset$PERSDOC3 <- replace(CDC_2023_subset$PERSDOC3, CDC_2023_subset$PERSDOC3 == 1 & CDC_2023_subset$PERSDOC3 == 2, 1) 
CDC_2023_subset$PERSDOC3 <- replace(CDC_2023_subset$PERSDOC3, CDC_2023_subset$PERSDOC3 == 3, 2)
CDC_2023_subset$PERSDOC3 <- as.factor(CDC_2023_subset$PERSDOC3)

# DRNK3GE5 will be simplified to indicate that a person has been binge drinking at least 1 time in last 30 days with 1 and no drinks will be indicated by 0 (originally indicated by 88)
CDC_2023_subset$DRNK3GE5 <- replace(CDC_2023_subset$DRNK3GE5, CDC_2023_subset$DRNK3GE5 == 88, 0)
CDC_2023_subset$DRNK3GE5 <- replace(CDC_2023_subset$DRNK3GE5, CDC_2023_subset$DRNK3GE5 >= 1, 1) 
CDC_2023_subset$DRNK3GE5 <- as.factor(CDC_2023_subset$DRNK3GE5)

# DIABETE4 will be simplified so 0 indicates the individual does not have diabetes (3) and 1 indicating the individual has type 2 diabetes or pre-diabetes or borderline diabetes (1 and 2 for DIABTYPE or 4)
CDC_2023_subset$DIABETE4 <- as.numeric(CDC_2023_subset$DIABETE4)
CDC_2023_subset$DIABETE4 <- replace(CDC_2023_subset$DIABETE4, CDC_2023_subset$DIABETE4 == 3, 0)
CDC_2023_subset$DIABETE4 <- replace(CDC_2023_subset$DIABETE4, CDC_2023_subset$DIABETE4 == 1 | CDC_2023_subset$DIABETE4 == 4, 1)
CDC_2023_subset$DIABETE4 <- as.factor(CDC_2023_subset$DIABETE4)

# BPHIGH6 will be simplified so 1 indicates one has had high or borderline high or pre-hypertensive or elevated blood pressure
CDC_2023_subset$BPHIGH6 <- as.numeric(CDC_2023_subset$BPHIGH6)
CDC_2023_subset$BPHIGH6 <- replace(CDC_2023_subset$BPHIGH6, CDC_2023_subset$BPHIGH6 == 2 | CDC_2023_subset$BPHIGH6 == 3, 2)
CDC_2023_subset$BPHIGH6 <- replace(CDC_2023_subset$BPHIGH6, CDC_2023_subset$BPHIGH6 == 1 | CDC_2023_subset$BPHIGH6 == 4, 1)
CDC_2023_subset$BPHIGH6 <- as.factor(CDC_2023_subset$BPHIGH6)
```

# Exploratory Data Analysis

## Libraries for data visualizations
```{r}
library(ggplot2)
library(dplyr)
library(scales)
library(ggbeeswarm)
library(treemap)
library(plotly)
```

## Get columns proportions for factor data after data cleaning

### Column proportion summaries
```{r}
CDC_2023_subset %>%
  tbl_summary()
```

### EDUCA
```{r}
# Proportion summary
table <- table(CDC_2023_subset$EDUCA, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=EDUCA, fill=EDUCA)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=EDUCA, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### EMPLOY1 (1:employed, 2: out of labor force, 3: unemployed)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$EMPLOY1, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=EMPLOY1, fill=EMPLOY1)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=EMPLOY1, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### INCOME3
```{r}
# Proportion summary
table <- table(CDC_2023_subset$INCOME3, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=INCOME3, fill=INCOME3)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=INCOME3, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### SEXVAR (1:male, 2:female)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$SEXVAR, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=SEXVAR, fill=SEXVAR)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=SEXVAR, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### PHYSHLTH (0: 0 days of bad health, 1: 1-13 days of bad health, 2: 14+ days of bad health)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$PHYSHLTH, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=PHYSHLTH, fill=PHYSHLTH)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=PHYSHLTH, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### MENTHLTH (0: 0 days of bad health, 1: 1-13 days of bad health, 2: 14+ days of bad health)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$MENTHLTH, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=MENTHLTH, fill=MENTHLTH)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=MENTHLTH, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### PERSDOC3 (1: 1 or more personal doctors, 2: none)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$PERSDOC3, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=PERSDOC3, fill=PERSDOC3)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=PERSDOC3, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### MEDCOST1 (1: med cost too high in last 12 months, 2: not)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$MEDCOST1, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=MEDCOST1, fill=MEDCOST1)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=MEDCOST1, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### SMOKE100 (1: smoked at least 100 cigarettes, 2: has not smoked more than 100 cigarettes)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$SMOKE100, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=SMOKE100, fill=SMOKE100)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=SMOKE100, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### DRNK3GE5 (0: No binge drinking in last 30 days, 1: yes, binge drinking in last 30 days)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$DRNK3GE5, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=DRNK3GE5, fill=DRNK3GE5)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=DRNK3GE5, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### EXERANY2 (1: yes,physical activity in last 30 days, 2: no physical activity in last 30 days)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$EXERANY2, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=EXERANY2, fill=EXERANY2)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=EXERANY2, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### BPHIGH6 (1: Told BP high, 2: BP high in pregnancy, 3: Not high, 4: borderline or pre-hypertenisve or elevated BP)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$BPHIGH6, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=BPHIGH6, fill=BPHIGH6)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=BPHIGH6, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0.25, size=3)
```

### TOLDHI3 (1: high cholesterol, 2: not high)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$TOLDHI3, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=TOLDHI3, fill=TOLDHI3)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=TOLDHI3, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0, size=3)
```

### CVDINFR4 (1: had heart attack, 2: no heart attack)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$CVDINFR4, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=CVDINFR4, fill=CVDINFR4)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=CVDINFR4, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0, size=3)
```

### CVDCRHD4 (1: angina or coronary heart disease, 2: no)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$CVDCRHD4, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=CVDCRHD4, fill=CVDCRHD4)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=CVDCRHD4, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0, size=3)
```

### CVDSTRK3 (1: had a stroke, 2: no)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$CVDSTRK3, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph overall
ggplot(CDC_2023_subset, aes(x=CVDSTRK3, fill=CVDSTRK3)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=CVDSTRK3, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0, size=3)
```

### CHCKDNY2 (1: had kidney disease, 2: no)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$CHCKDNY2, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph
ggplot(CDC_2023_subset, aes(x=CHCKDNY2, fill=CHCKDNY2)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)

# Bar graph by diabetes
CDC_2023_subset %>% 
  group_by(DIABETE4) %>%
  ggplot(aes(x=CHCKDNY2, fill=DIABETE4)) + 
    geom_bar() + 
    scale_y_continuous(labels = scales::comma) +
    geom_text(stat='count', aes(label=..count..), vjust = 0, size=3)
```

### Diabetes (0: does not have diabetes, 1: has type 2 diabetes or prediabetes or borderline diabetes)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$DIABETE4, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph
ggplot(CDC_2023_subset, aes(x=DIABETE4, fill=DIABETE4)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)
```

### DIABTYPE summary (2: type 2 diabetes, NA: no diabetes)
```{r}
# Proportion summary
table <- table(CDC_2023_subset$DIABTYPE, useNA = "ifany")
prop.table(table) %>% {. * 100} %>% round(4)

# Bar graph
ggplot(CDC_2023_subset, aes(x=DIABTYPE, fill=DIABTYPE)) + 
  geom_bar() + 
  scale_y_continuous(labels = scales::comma)
```

## Compare Diabetes and Exercise to Determine the Impact on Type 2 Diabetes
```{r}
# Convert DIABETE4 and EXERANY2 to correct types
CDC_2023_subset$EXERANY2 <- as.factor(CDC_2023_subset$EXERANY2)
CDC_2023_subset$DIABETE4 <- as.numeric(as.character(CDC_2023_subset$DIABETE4))

# Filter out inappropriate data
CDC_2023_subset <- CDC_2023_subset %>%
  filter(EXERANY2 %in% c("1", "2"), !is.na(DIABETE4), !is.na(BMI))

# Categorize BMI into groups
CDC_2023_subset <- CDC_2023_subset %>%
  mutate(BMI_Category = case_when(
    BMI < 18.5 ~ "Underweight",
    BMI >= 18.5 & BMI < 25 ~ "Normal weight",
    BMI >= 25 & BMI < 30 ~ "Overweight",
    BMI >= 30 ~ "Obese"
  ))

# Calculate the prevalence of diabetes by exercise and BMI category
exercise_bmi_diabetes_data <- CDC_2023_subset %>%
  group_by(EXERANY2, BMI_Category) %>%
  summarise(Diabetes_Prevalence = mean(DIABETE4 == 1, na.rm = TRUE),
            .groups = 'drop') %>%
  mutate(Exercise = ifelse(EXERANY2 == "1", "Yes", "No"))

# Create the bar plot with BMI categories
ggplot(exercise_bmi_diabetes_data, aes(x = Exercise, y = Diabetes_Prevalence, fill = BMI_Category)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Prevalence of Diabetes by Exercise and BMI Category",
       x = "Regular Exercise",
      
)
```


## Measures of central tendencies, spread, and measures of distribution shape (skew and kurtosis) and histogram for all and type 2 diabetes
```{r}
# Summary of BMI column
library(moments)
print(summary(CDC_2023_subset$BMI))
cat("Skewness is ", skewness(CDC_2023_subset$BMI), ". ")
cat("Kurtosis is ", kurtosis(CDC_2023_subset$BMI), ".")

# Histogram 
BMI_summ1 <- CDC_2023_subset %>% summarize(mean = mean(BMI), median = median(BMI)) #Calculate mean for graph
BMI_over_75 <- filter(CDC_2023_subset, BMI>75)
outliers <- data.frame(x = BMI_over_75$BMI, y = 0)
ggplot(data = CDC_2023_subset, aes(x = BMI)) +
  geom_histogram(aes(y= ..density..), binwidth = 0.5, fill="grey") +
  geom_density(linewidth = 0.7) +
  geom_vline(aes(xintercept = mean), BMI_summ1, color = "red", linewidth = 0.7) +
  geom_vline(aes(xintercept = median), BMI_summ1, color = "blue", linewidth = 0.7) + 
  xlab("BMI") +
  ylab("Density") + 
  labs(title="Density Plot of BMI") + 
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(data = outliers, aes(x, y), height = 0.1) +
  scale_x_continuous(breaks=c(0,30,60,90,120)) +
  xlim(0,120) 

BMI_summ2 <- CDC_2023_subset %>% filter(Type_2_Diabetes == 1) %>% summarize(mean = mean(BMI), median = median(BMI)) #Calculate mean for graph

BMI_diab <- CDC_2023_subset %>%
  filter(Type_2_Diabetes == 1) 
ggplot(data=BMI_diab, aes(x = BMI)) +
    geom_histogram(aes(y= ..density..), binwidth = 0.5, fill="grey") +
    geom_density(linewidth = 0.7) +
    geom_vline(aes(xintercept = mean), BMI_summ2, color = "red", linewidth = 0.7) +
    geom_vline(aes(xintercept = median), BMI_summ2, color = "blue", linewidth = 0.7) + 
    xlab("BMI") +
    ylab("Density") + 
    labs(title="Density Plot of BMI for Type 2 Diabetics") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0,30,60,90,120)) +
    xlim(0,120) 
```

## Proportion tables split by social factors, demographic characteristics, and health conditions

### Change column names for simpler axes and legends
```{r}
colnames(CDC_2023_subset) <- c("Education", "Employment_Status", "Income", "Weight", "Height", "Sex", "Physical_Health", "Mental_Health", "Personal_Doctor", "Not_Afford_Doctor", "Smoked_100_Pkgs", "Binge_Drinking", "Exercised", "HighBP", "High_Cholesterol", "Myocardial_Infarction", "Heart_Disease", "Stroke", "Kidney_Disease", "Type_2_Diabetes", "DIABTYPE", "NA", "BMI")
```

### Tables for each of our categories (Social Factors, Demographics, and Diseases)
```{r}
# Social factors
CDC_2023_subset %>%
  mutate(Physical_Health = str_replace(Physical_Health, "0", "0 Days")) %>%
  mutate(Physical_Health = str_replace(Physical_Health, "1", "1 -13 Days")) %>%
  mutate(Physical_Health = str_replace(Physical_Health, "2", "14+ Days")) %>%
  mutate(Mental_Health = str_replace(Mental_Health, "0", "0 Days")) %>%
  mutate(Mental_Health = str_replace(Mental_Health, "1", "1 -13 Days")) %>%
  mutate(Mental_Health = str_replace(Mental_Health, "2", "14+ Days")) %>%
  mutate(Personal_Doctor = str_replace(Personal_Doctor, "1", "More than one")) %>%
  mutate(Personal_Doctor = str_replace(Personal_Doctor, "2", "No")) %>%
  mutate(Not_Afford_Doctor = str_replace(Not_Afford_Doctor, "1", "Yes")) %>%
  mutate(Not_Afford_Doctor = str_replace(Not_Afford_Doctor, "2", "No")) %>%
  mutate(Smoked_100_Pkgs = str_replace(Smoked_100_Pkgs, "1", "Yes")) %>%
  mutate(Smoked_100_Pkgs = str_replace(Smoked_100_Pkgs, "2", "No")) %>%
  mutate(Binge_Drinking = str_replace(Binge_Drinking, "1", "Yes")) %>%
  mutate(Binge_Drinking = str_replace(Binge_Drinking, "0", "No")) %>%
  mutate(Exercised = str_replace(Exercised, "1", "Yes")) %>%
  mutate(Exercised = str_replace(Exercised, "2", "No")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  tbl_summary(
    include = c(Physical_Health, Mental_Health, Personal_Doctor, Not_Afford_Doctor, Smoked_100_Pkgs, Binge_Drinking, Exercised),
    by = Type_2_Diabetes,
    missing = 'no',
    type = all_dichotomous() ~ "categorical"
  ) |>
  add_n() |>
  modify_header()

# Demographic characteristics
CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "1", "Employed")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "2", "Out of Labor Force")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "3", "Unemployed")) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  tbl_summary(
    include = c(Education, Employment_Status, Income, Sex, BMI),
    by = Type_2_Diabetes,
    missing = 'no',
    type = all_dichotomous() ~ "categorical"
  ) |>
  add_n() |>
  modify_header()

# Health conditions
CDC_2023_subset %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(HighBP = str_replace(HighBP, "1", "Yes")) %>%
  mutate(HighBP = str_replace(HighBP, "2", "No")) %>%
  mutate(High_Cholesterol = str_replace(High_Cholesterol, "1", "Yes")) %>%
  mutate(High_Cholesterol = str_replace(High_Cholesterol, "2", "No")) %>%
  mutate(Myocardial_Infarction = str_replace(Myocardial_Infarction, "1", "Yes")) %>%
  mutate(Myocardial_Infarction = str_replace(Myocardial_Infarction, "2", "No")) %>%
  mutate(Heart_Disease = str_replace(Heart_Disease, "1", "Yes")) %>%
  mutate(Heart_Disease = str_replace(Heart_Disease, "2", "No")) %>%
  mutate(Stroke = str_replace(Stroke, "1", "Yes")) %>%
  mutate(Stroke = str_replace(Stroke, "2", "No")) %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "1", "Yes")) %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "2", "No")) %>%
  tbl_summary(
    include = c(HighBP, High_Cholesterol, Myocardial_Infarction, Heart_Disease, Stroke, Kidney_Disease),
    by = Type_2_Diabetes,
    type = all_dichotomous() ~ "categorical"
  ) |>
  add_n() |>
  modify_header()
```

### Comparing BMI with Education, Income, Employment, Ability to afford doctor, and Type 2 diabetes
```{r}
CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot() +
    aes(y = BMI, x = Education) +
    geom_quasirandom(aes()) +
    ylab("BMI") +
    xlab("Education Level") + 
    labs(title="Impact of Education on BMI") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```
```{r}
CDC_2023_subset %>%
  mutate(Income = str_replace(Income, "1", "Less than $10,000")) %>%
  mutate(Income = str_replace(Income, "2", "$10,000 to <$15,000")) %>%
  mutate(Income = str_replace(Income, "3", "$15,000 to <$20,000")) %>%
  mutate(Income = str_replace(Income, "4", "$20,000 to <$25,000")) %>%
  mutate(Income = str_replace(Income, "5", "$25,000 to <$35,000")) %>%
  mutate(Income = str_replace(Income, "6", "$35,000 to <$50,000")) %>%
  mutate(Income = str_replace(Income, "7", "$50,000 to <$75,000")) %>%
  mutate(Income = str_replace(Income, "8", "$75,000 to < $100,000")) %>%
  mutate(Income = str_replace(Income, "9", "$100,000 to <$150,000")) %>%
  mutate(Income = str_replace(Income, "10", "$150,000 to <$200,000")) %>%
  mutate(Income = str_replace(Income, "11", "$200,000 or more")) %>%
  mutate(Income = as.factor(Income)) %>%
  ggplot(aes(y = BMI, x = Income)) +
    geom_quasirandom(aes(y = BMI, x = Income)) +
    ylab("BMI") +
    xlab("Income Level") + 
    labs(title="Impact of Income Level on BMI") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Employment_Status = str_replace(Employment_Status, "1", "Employed")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "2", "Out of Labor Force")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "3", "Unemployed")) %>%
  ggplot(aes(y = BMI, x = Employment_Status)) +
    geom_quasirandom(aes(y = BMI, x = Employment_Status)) +
    ylab("BMI") +
    xlab("Employment Status") + 
    labs(title="Impact of Employment Status on BMI") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Not_Afford_Doctor = str_replace(Not_Afford_Doctor, "1", "Yes")) %>%
  mutate(Not_Afford_Doctor = str_replace(Not_Afford_Doctor, "2", "No")) %>%
  ggplot(aes(y = BMI, x = Not_Afford_Doctor)) +
    geom_quasirandom(aes()) +
    ylab("Was individual unable to afford a doctor in the last 12 months?") +
    xlab("BMI") + 
    labs(title="Impact of Ability to Afford Doctor on BMI") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  ggplot() +
    aes(y = BMI, x = Type_2_Diabetes) +
    geom_quasirandom(aes()) +
    xlab("Does the individual have type 2 diabetes?") +
    ylab("BMI") + 
    labs(title="Relationship Between Type 2 Diabetes and BMI") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

#### Education compared with Diseases
```{r}
CDC_2023_subset %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = Type_2_Diabetes, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education Level") + 
    labs(title="Education Level Distribution and Proportion of Type 2 Diabetes") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = Type_2_Diabetes, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education Level") + 
    labs(title="Education Level Distribution and Proportion of Type 2 Diabetes") + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
library(ggtext)
CDC_2023_subset %>%
  mutate(HighBP = str_replace(HighBP, "1", "Yes")) %>%
  mutate(HighBP = str_replace(HighBP, "2", "No")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = HighBP, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with High Blood Pressure")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(HighBP = str_replace(HighBP, "1", "Yes")) %>%
  mutate(HighBP = str_replace(HighBP, "2", "No")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = HighBP, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with High Blood Pressure")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(High_Cholesterol = str_replace(High_Cholesterol, "1", "Yes")) %>%
  mutate(High_Cholesterol = str_replace(High_Cholesterol, "2", "No")) %>%
  ggplot(aes(fill = High_Cholesterol, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with High Cholesterol")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(High_Cholesterol = str_replace(High_Cholesterol, "1", "Yes")) %>%
  mutate(High_Cholesterol = str_replace(High_Cholesterol, "2", "No")) %>%
  ggplot(aes(fill = High_Cholesterol, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with High Cholesterol")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Myocardial_Infarction = str_replace(Myocardial_Infarction, "1", "Yes")) %>%
  mutate(Myocardial_Infarction = str_replace(Myocardial_Infarction, "2", "No")) %>%
  ggplot(aes(fill = Myocardial_Infarction, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Myocardial Infarction")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Myocardial_Infarction = str_replace(Myocardial_Infarction, "1", "Yes")) %>%
  mutate(Myocardial_Infarction = str_replace(Myocardial_Infarction, "2", "No")) %>%
  ggplot(aes(fill = Myocardial_Infarction, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Myocardial Infarction")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Heart_Disease = str_replace(Heart_Disease, "1", "Yes")) %>%
  mutate(Heart_Disease = str_replace(Heart_Disease, "2", "No")) %>%
  ggplot(aes(fill = Heart_Disease, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Heart Disease")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Heart_Disease = str_replace(Heart_Disease, "1", "Yes")) %>%
  mutate(Heart_Disease = str_replace(Heart_Disease, "2", "No")) %>%
  ggplot(aes(fill = Heart_Disease, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Heart Disease")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Stroke = str_replace(Stroke, "1", "Yes")) %>%
  mutate(Stroke = str_replace(Stroke, "2", "No")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = Stroke, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Stroke")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(Stroke = str_replace(Stroke, "1", "Yes")) %>%
  mutate(Stroke = str_replace(Stroke, "2", "No")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = Stroke, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Stroke")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

```{r}
CDC_2023_subset %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "1", "Yes")) %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "2", "No")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = Kidney_Disease, x = Education)) + 
    geom_bar(position="stack") +
    ylab("Count") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Kidney Disease")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()

CDC_2023_subset %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "1", "Yes")) %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "2", "No")) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  ggplot(aes(fill = Kidney_Disease, x = Education)) + 
    geom_bar(position="fill") +
    ylab("Proportion") +
    xlab("Education") + 
    labs(title=str_wrap("Education Level Distribution and Proportion of Individuals with Kidney Disease")) + 
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```

### Heatmap for Correlation

#### Disease
```{r}
library(ggcorrplot)
subset <- CDC_2023_subset %>%
  select(HighBP, High_Cholesterol, Myocardial_Infarction, Heart_Disease, Stroke, Kidney_Disease, Type_2_Diabetes) 
model.matrix(~0+., data=subset) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=TRUE, lab_size=2)
```

#### Social Factors
```{r}
subset <- CDC_2023_subset %>%
  select(Physical_Health, Mental_Health, Personal_Doctor, Not_Afford_Doctor, Smoked_100_Pkgs, Binge_Drinking, Exercised) 
model.matrix(~0+., data=subset) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=TRUE, lab_size=2)
```

#### Demographics
```{r}
subset <- CDC_2023_subset %>%
  select(Education, Employment_Status, Income, Sex, BMI) 
model.matrix(~0+., data=subset) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=TRUE, lab_size=2)
```

### Plots to compare relationship between Demographic Variables
#### Different size dots for different counts (use to see if there is a relationship between the two variables)
```{r}
df_summary <- CDC_2023_subset %>%                                                                 
  group_by(Education, Employment_Status, Sex) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "1", "Employed")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "2", "Out of Labor Force")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "3", "Unemployed")) %>%
  mutate(Income = str_replace(Income, "1", "Less than $10,000")) %>%
  mutate(Income = str_replace(Income, "2", "$10,000 to <$15,000")) %>%
  mutate(Income = str_replace(Income, "3", "$15,000 to <$20,000")) %>%
  mutate(Income = str_replace(Income, "4", "$20,000 to <$25,000")) %>%
  mutate(Income = str_replace(Income, "5", "$25,000 to <$35,000")) %>%
  mutate(Income = str_replace(Income, "6", "$35,000 to <$50,000")) %>%
  mutate(Income = str_replace(Income, "7", "$50,000 to <$75,000")) %>%
  mutate(Income = str_replace(Income, "8", "$75,000 to < $100,000")) %>%
  mutate(Income = str_replace(Income, "9", "$100,000 to <$150,000")) %>%
  mutate(Income = str_replace(Income, "10", "$150,000 to <$200,000")) %>%
  mutate(Income = str_replace(Income, "11", "$200,000 or more")) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Count = n()) 
ggplot(df_summary, aes(x = Education, y = Employment_Status, size=Count, color=Count)) +             
  geom_point() +
  facet_wrap(~Sex) +
  coord_flip() + 
  ylab("Employment Status") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
df_summary <- CDC_2023_subset %>%                                                                 
  group_by(Education, Income, Sex) %>%
  mutate(Count = n()) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "1", "Employed")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "2", "Out of Labor Force")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "3", "Unemployed")) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  summarise(Count = n())
  
ggplot(df_summary, aes(x = Education, y = Income, size=Count, color=Count)) +             
  geom_point() +
  facet_wrap(~Sex) +
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Plots to compare relationship between Demographic and Social Factors Variables
#### Different size dots for different counts (use to see if there is a relationship between the two variables)
```{r}
df_summary <- CDC_2023_subset %>%                                                                 
  group_by(Education, Mental_Health, Sex) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Mental_Health = str_replace(Mental_Health, "0", "0 days")) %>%
  mutate(Mental_Health = str_replace(Mental_Health, "1", "0-13 days")) %>%
  mutate(Mental_Health = str_replace(Mental_Health, "2", "14+ days")) %>%
  mutate(Mental_Health = factor(Mental_Health, levels=c("0 days", "0-13 days", "14+ days"))) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Count = n())
ggplot(df_summary, aes(x = Education, y = Mental_Health, size=Count, color=Count)) +             
  geom_point() +
  facet_wrap(~Sex) +
  coord_flip() + 
  ylab("Days with Poor Mental Health") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r}
df_summary <- CDC_2023_subset %>%                                                                 
  group_by(Education, Not_Afford_Doctor, Sex) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Not_Afford_Doctor = str_replace(Not_Afford_Doctor, "1", "Yes")) %>%
  mutate(Not_Afford_Doctor = str_replace(Not_Afford_Doctor, "2", "No")) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Count = n())
ggplot(df_summary, aes(x = Education, y = Not_Afford_Doctor, size=Count, color=Count)) +             
  geom_point() +
  facet_wrap(~Sex) +
  coord_flip() + 
  ylab("Days with Poor Mental Health") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
df_summary <- CDC_2023_subset %>%                                                                 
  group_by(Education, Physical_Health, Sex) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Physical_Health = str_replace(Physical_Health, "0", "0 days")) %>%
  mutate(Physical_Health = str_replace(Physical_Health, "1", "0-13 days")) %>%
  mutate(Physical_Health = str_replace(Physical_Health, "2", "14+ days")) %>%
  mutate(Physical_Health = factor(Physical_Health, levels=c("0 days", "0-13 days", "14+ days"))) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Count = n())
ggplot(df_summary, aes(x = Education, y = Physical_Health, size=Count, color=Count)) +             
  geom_point() +
  facet_wrap(~Sex) +
  coord_flip() + 
  ylab("Days with Poor Physical Health") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Plots to compare relationship between Type 2 Diabetes and Kidney Disease
#### Different size dots for different counts (use to see if there is a relationship between the two variables)
```{r}
df_summary <- CDC_2023_subset %>%                                                                 
  group_by(Type_2_Diabetes, Kidney_Disease, Sex) %>%
  mutate(Education = str_replace(Education, "1", "Never attended school or only attended kindergarten")) %>%
  mutate(Education = str_replace(Education, "2", "Grades 1 through 8 (Elementary)")) %>%
  mutate(Education = str_replace(Education, "3", "Grades 9 through 11 (Some high school)")) %>%
  mutate(Education = str_replace(Education, "4", "Grade 12 or GED (High school graduate)")) %>%
  mutate(Education = str_replace(Education, "5", "College 1 to 3 years (Some college or technical school)")) %>%
  mutate(Education = str_replace(Education, "6", "College 4 years or more (College graduate)")) %>%
  mutate(Education = factor(Education, levels=c("Never attended school or only attended kindergarten", "Grades 1 through 8 (Elementary)", "Grades 9 through 11 (Some high school)", "Grade 12 or GED (High school graduate)", "College 1 to 3 years (Some college or technical school)", "College 4 years or more (College graduate)"))) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "1", "Employed")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "2", "Out of Labor Force")) %>%
  mutate(Employment_Status = str_replace(Employment_Status, "3", "Unemployed")) %>%
  mutate(Income = str_replace(Income, "1", "Less than $10,000")) %>%
  mutate(Income = str_replace(Income, "2", "$10,000 to <$15,000")) %>%
  mutate(Income = str_replace(Income, "3", "$15,000 to <$20,000")) %>%
  mutate(Income = str_replace(Income, "4", "$20,000 to <$25,000")) %>%
  mutate(Income = str_replace(Income, "5", "$25,000 to <$35,000")) %>%
  mutate(Income = str_replace(Income, "6", "$35,000 to <$50,000")) %>%
  mutate(Income = str_replace(Income, "7", "$50,000 to <$75,000")) %>%
  mutate(Income = str_replace(Income, "8", "$75,000 to < $100,000")) %>%
  mutate(Income = str_replace(Income, "9", "$100,000 to <$150,000")) %>%
  mutate(Income = str_replace(Income, "10", "$150,000 to <$200,000")) %>%
  mutate(Income = str_replace(Income, "11", "$200,000 or more")) %>%
  mutate(Sex = str_replace(Sex, "1", "Male")) %>%
  mutate(Sex = str_replace(Sex, "2", "Female")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "0", "No Diabetes")) %>%
  mutate(Type_2_Diabetes = str_replace(Type_2_Diabetes, "1", "Type 2 Diabetes")) %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "1", "Yes")) %>%
  mutate(Kidney_Disease = str_replace(Kidney_Disease, "2", "No")) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
ggplot(df_summary, aes(x = Type_2_Diabetes, y = Kidney_Disease, size=freq, color=freq)) +             
  geom_point() +
  facet_wrap(~Sex) +
  coord_flip() + 
  xlab("Type 2 Diabetes") +
  ylab("Kidney Disease") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```